Two Stars and Two Signals in PTFO 8-8695

### 2020-04-26 paper/ms.tex
minor
call an ice-cream truck
we've been scooped! not the end of the world, though.

%\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}
% \title{Seeing Double: Still Against the Planetary Interpretation of
% PTFO$\,$8-8695b}
% \title{Seeing Double: TESS and Gaia Show That PTFO$\,$8-8695b is Unlikely a
% Planet}
%\title{Seeing Double: Indications From {\it TESS} and {\it Gaia} That
%PTFO$\,$8-8695b is Not a Planet}
%\title{Double Trouble: Transient Dips and Photometric Binarity in
%PTFO$\,$8-8695}
\title{Double, Double, Spin plus Nubble: Two Stars and Two Signals in
PTFO 8-8695}

@@

%
% key authors:
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}

@@

\epigraph{
	Nubble [n]: a small lump or protuberance.
}

@@

These lightcurves have been phenomenologically classified as
``persistent flux dips'' or ``transient flux dips'', based on whether
their depths and durations show variability over the 90-day K2
campaigns \citep{stauffer_orbiting_2017}.
As \citet{stauffer_orbiting_2017} explain,
these  objects are morphologically distinct from ``scallop shell' lightcurves,
and are present in stars at more advanced evolutionary disk stages than
the ``dipper'' stars \citep{ansdell_young_2016,cody_manyfaceted_2018}.
The persistent and transient flux dip stars all show angular dips
that are cannot be explained with starspots.
They typically have the following properties.
\begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
  \item Weak-lined T Tauri stars.
  \item Spectral type typically M2-M5 ({\it e.g.},
    \citealt{rebull_usco_2018},~Figure~20).
  \item Ages $\lesssim$ 100$\,$Myr.
  \item Stars typically lack spectroscopic accretion indicators 
  \item Phased lightcurves show shallow, angular dips, usually
    superposed on large-amplitude smooth variability. The latter
    is interpreted as stellar rotation.
  \item Rotation is fast---usually between 0.5 and 2.0 days.
  \item Rarely any infrared excess (never any W4 detection; only a few
    cases with W3 detctions).
  \item Can show multiple dips per cycle.
  %({\it e.g.}, EPIC 203692610, RIK-210).
  \item Dip depths, durations, and phases can vary over just a few cycles
    ({\it e.g.}, EPIC 204143627).
  \item Dip depths can change after flares.
  \item Rare at a population level: $\lesssim 1\%$ of relevant stars
    \citep{rebull_usco_2018}.
\end{enumerate}
The 10.74 hour signal in \ptfo\ meets all of these criteria.
This is the first connection of \ptfo\ with this class of objects
because the TESS lightcurve was needed to resolve the different
rotation signals.

@@

\acknowledgements
%
%This paper includes data collected by the TESS mission, which are
%publicly available from the Mikulski Archive for Space Telescopes
%(MAST).
%
%Funding for the TESS mission is provided by NASA's Science Mission
%directorate.
%
The authors thank D.~Fabrycky for helpful tidal evolution calculations
that were eventually omitted.
%
\ptfo\ was included on the TESS ``short-cadence'' target list thanks to the
Guest Investigator programs of S.\ Czesla and C.\ Huang (G011128 and
G011132 respectively).
%
The Digitized Sky Survey was produced at the Space Telescope Science
Institute under U.S. Government grant NAG W-2166.
Figure~\ref{fig:scene} is based on photographic data obtained using
the Oschin Schmidt Telescope on Palomar Mountain.
%
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% (Some of) The data presented herein were obtained at the W. M. Keck
% Observatory, which is operated as a scientific partnership among the
% California Institute of Technology, the University of California and
% the National Aeronautics and Space Administration. The Observatory was
% made possible by the generous financial support of the W. M. Keck
% Foundation.
% The authors wish to recognize and acknowledge the very significant
% cultural role and reverence that the summit of Maunakea has always had
% within the indigenous Hawaiian community.  We are most fortunate to
% have the opportunity to conduct observations from this mountain.
%
% \newline
%

>>>>

%\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}
% \title{Seeing Double: Still Against the Planetary Interpretation of
% PTFO$\,$8-8695b}
% \title{Seeing Double: TESS and Gaia Show That PTFO$\,$8-8695b is Unlikely a
% Planet}
%\title{Seeing Double: Indications From {\it TESS} and {\it Gaia} That
%PTFO$\,$8-8695b is Not a Planet}
%\title{Double Trouble: Transient Dips and Photometric Binarity in
%PTFO$\,$8-8695}
%\title{Double, Double, Spin plus Nubble: Two Stars and Two Signals in
%PTFO 8-8695}
\title{Two Stars and Two Signals in PTFO 8-8695}

@@

%
% key authors:
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-4265-047X]{...others!}
\affiliation{ Department of Astrophysical Sciences, Princeton
	University, 4 Ivy Lane, Princeton, NJ 08540, USA}

@@

%\epigraph{
%	Nubble [n]: a small lump or protuberance.
%}

@@

These lightcurves have been phenomenologically classified as
``persistent flux dips'' or ``transient flux dips'', based on whether
their depths and durations show variability over the 90-day K2
campaigns \citep{stauffer_orbiting_2017}.
As \citet{stauffer_orbiting_2017} explain,
these  objects are morphologically distinct from ``scallop shell' lightcurves,
and are present in stars at more advanced evolutionary disk stages than
the ``dipper'' stars \citep{ansdell_young_2016,cody_manyfaceted_2018}.
The persistent and transient flux dip stars all show angular dips
that are cannot be explained with starspots.
They typically have the following properties.
\begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
  \item Weak-lined T Tauri stars.
    (Few to no spectroscopic accretion indicators). 
  \item Spectral type typically M2-M5 ({\it e.g.},
    \citealt{rebull_usco_2018},~Figure~20).
  \item Ages $\lesssim$ 100$\,$Myr.
  \item Phased lightcurves show shallow, angular dips, usually
    superposed on large-amplitude smooth variability. The latter
    is interpreted as stellar rotation.
  \item Rotation is fast---usually between 0.5 and 2.0 days.
  \item Rarely any infrared excess (never any W4 detection; only a few
    cases with W3 detctions).
  \item Can show multiple dips per cycle.
  %({\it e.g.}, EPIC 203692610, RIK-210).
  \item Dip depths, durations, and phases can vary over just a few cycles
    ({\it e.g.}, EPIC 204143627).
  \item Dip depths can change after flares.
  \item Rare at a population level: $\lesssim 1\%$ of relevant stars
    \citep{rebull_usco_2018}.
\end{enumerate}
The 10.74 hour signal in \ptfo\ meets all of these criteria.
This is the first connection of \ptfo\ with this class of objects
because the TESS lightcurve was needed to resolve the different
rotation signals.

@@

\acknowledgements
When this manuscript was at an advanced stage, we learned of the
complementary work by \citet{koen_2020}, which was in press at the
{\it Monthly Notices} before submission of our manuscript.  Our
studies independently reached the key conclusions that the TESS
lightcurve shows two periodic signals, and that the properties of
\ptfo\ are consistent with the emerging class of transient and
persistent flux dip stars.  \citet{koen_2020} reached these
conclusions by modeling the TESS lightcurve as a truncated sum of
Fourier terms, and concluded that the two signals are most simply
interpreted as coming from two stars.  Our analysis of the Gaia data
provide independent support for the conclusion that \ptfo\ is a
binary.

Minor differences in our approaches are that \citet{koen_2020}
interprets the ``out of dip'' signal to potentially show multiple
dips, while our BIC-based model comparison did not indicate the need
for additional harmonics after fitting for the largest dip and two
harmonics at $P_{\rm s}$ (Table~1). Our analysis emphasized the
agreement between the TESS dip ephemeris and that from
\citet{tanimoto_evidence_2020}.  \citet{koen_2020} however made a
clearer case for testing the binarity hypothesis using interferometric
imaging and high-precision radial velocities.
\\
\\
%
%This paper includes data collected by the TESS mission, which are
%publicly available from the Mikulski Archive for Space Telescopes
%(MAST).
%
%Funding for the TESS mission is provided by NASA's Science Mission
%directorate.
%
The authors thank D.~Fabrycky for helpful tidal evolution calculations
that were eventually omitted.
%
\ptfo\ was included on the TESS ``short-cadence'' target list thanks to the
Guest Investigator programs of S.\ Czesla and C.\ Huang (G011128 and
G011132 respectively).
%
The Digitized Sky Survey was produced at the Space Telescope Science
Institute under U.S. Government grant NAG W-2166.
Figure~\ref{fig:scene} is based on photographic data obtained using
the Oschin Schmidt Telescope on Palomar Mountain.
%
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% (Some of) The data presented herein were obtained at the W. M. Keck
% Observatory, which is operated as a scientific partnership among the
% California Institute of Technology, the University of California and
% the National Aeronautics and Space Administration. The Observatory was
% made possible by the generous financial support of the W. M. Keck
% Foundation.
% The authors wish to recognize and acknowledge the very significant
% cultural role and reverence that the summit of Maunakea has always had
% within the indigenous Hawaiian community.  We are most fortunate to
% have the opportunity to conduct observations from this mountain.
%
% \newline
%
